### 2018-02-20 submission/manuscript.Rmd
update results section

***The Number of OTUs is Dependent on the HiFi DNA Polymerase Used:***
In order to compare the number of OTUs across individuals a Z-score normalization by individual, by cycle number, was employed on the number of OTUs data. After normalization, we identified that there was a HiFi DNA polymerase dependent difference that was consistent across sub-sampling [Figure 1]. Lower cycle numbers (15-20) tended to result in less differences between HiFI DNA polymerase while cycle numbers of 25, 30, and 35 tended to have more distinct differences [Figure 1]. All sub-sampling levels were significantly different only at 35 cycles (P-value `r ifelse(max(filter(fs_anova_summary, cycle == "35x")[, "bh"]) < 0.0001, invisible("<"), invisible("check data"))` 0.0001) [Table S1]. There were differences in HiFi DNA polymerase at 25 and 30 cycles but the sub-sampling depth had to be 5000 or higher (P-value `r ifelse(max(filter(fs_anova_summary, cycle == "30x", sub_sample_level != 1000)[, "bh"]) < 0.05 & max(filter(fs_anova_summary, cycle == "25x", sub_sample_level != 1000)[, "bh"]) < 0.05, invisible("<"), invisible("check data"))` 0.05) [Table S1]. Platinum HiFi DNA polymerase was the main driver of the differences observed across all sub-sampling depths at 35 cycles, based on a Tukey post-hoc test (P-value `r ifelse(max(filter(fs_tukey_summary, padj < 0.05)[, "padj"]) < 0.05,  invisible("<"), invisible("check data"))` 0.05) [Table S2].

This HiFi DNA polymerase dependent difference in the number of OTUs was also observed in the mock community samples with the same DNA polymerases having high (Platinum) and low (Accuprime) number of OTUs [Figure 2 & Table S3]. Conversely, differences between HiFi DNA polymerase were observed as early as 20 cycles and a sub-sampling depth of 1000 sequences (P-value = `r format(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "bh"], digits = 1, nsmall = 1)`) [Table S3]. Using a Tukey post-hoc test differences between Platinum and the other HiFi DNA polymerases was the major driver of the differences seen at different cycle numbers and sub-sampling depths [Table S4]. Both fecal and mock samples consistently showed that across sub-sampling depth and cycle number the lowest number of OTUs identified was from Accuprime^TM^ while the highest was from Platinum for both fecal and mock samples [Figure 1 & 2].

***Minimal Bray-Curtis Differences are Detected by Cycle Number:***
For both fecal and mock samples, there was little difference between each respective 5-cycle increment at higher cycle numbers and this was consistent across the different sub-samplings [Figure 3]. Overall, there were only two differences between 5-cycle increments that were identified. First, there were differences for the same fecal sample between 20 versus 25 cycles that was independent of HiFi DNA polymerase but dependent on sub-sampling depth (sub-sampled to 1000 = `r round(median(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`) (median (25% - 75% quantile)), sub-sampled to 5000 = `r round(median(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`), sub-sampled to 10000 = `r round(median(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`)) [Figure 3A]. Second, for the mocks, where data is available, there were larger difference between 20 and 25 cycles (sub-sampled to 1000 = `r round(median(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`)) [Figure 3B]. However, these differences between the next 5-cycle increment do not persist once 25 cycles are reached [Figure 3].

Using PERMANOVA to test for differences within HiFi DNA polymerase groups based on cycle number, only Phusion had cycle dependent differences at 1000 and 5000 sub-sampling depth (P-value = `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 1000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)` and `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 5000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)`, respectively). Phusion was also one of only two HiFi DNA polymerases that managed to have samples for the 1000 sub-sampling depth at 15 cycles. Next, we assessed whether there were any major differences between 5-cycle increments within each sample. We found that there was no detectable difference in Bray-Curtis index when comparing to the previous 5-cycle increment (P-value `r ifelse(min(bray_sim_to_prev_cycle_kruskal_results$bh, na.rm = T) > 0.05, invisible(">"), invisible("check data"))` 0.05). However, Phusion at 1000 sub-sampling depth had a P-value = `r format(filter(bray_sim_to_prev_cycle_kruskal_results, sub_sample_level == 1000, taq == "PHU")[, "pvalue"], digits = 1, nsmall = 1)` before multiple comparison correction. It should be noted that at higher sub-sampling depths these differences in Bray-Curtis indices disappear for both differences in cycle number and within 5-cycle increments within an individual.

***Sequence Error is Dependent on both HiFi DNA Polymerase and Cycle Number:***
Differences by HiFi DNA polymerase in the median average per base error varied without a clear pattern across sub-sampling depth [Table S5]. Generally, the highest per base median average error rates were for the Kappa HiFi DNA polymerase [Figure 4]. This error rate was minimally affected by both the 'pre.cluster' step and chimera removal by VSEARCH [Figure 4]. There were small differences in the per base error rate between the various HiFi DNA polymerase at lower cycle numbers and larger differences at higher cycle number with Platinum having the largest differences of all the HiFi DNA polymerase [Figure 4B-C and Table S6].

The total sequences with at least one error had multiple differences at different cycle numbers and was mostly alleviated by the use of the 'pre.cluster' step [Figure S1]. Major differences before this 'pre.cluster' step were driven by large differences in Accuprime^TM^ and Platinum versus the other HiFi DNA polymerase tested [Figure S1 & Table S7 & S8]. Although Accuprime^TM^ had the lowest per base error rate it had the largest number of sequences with at least one error, regardless of cycle number or sub-sampling depth [Figure S1]. However, this increased number of sequences with an error can be drastically lowered with existing bioinformatic approaches [Figure S1]. Investigation of whether there were HiFi DNA polymerase dependent effects on base substitution found that there were generally no biases in the types of substitution made [Figure S2].

***Chimeric Sequences are HiFi DNA Polymerase Dependent and Correlate with Number of OTUs:***
After chimera removal using VSEARCH and removal of sequences that did not classify as bacteria, we assessed the percentage of sequences that were still chimeric within our mock community. At all levels of sub-sampling and cycle number there were significant differences between the HiFi DNA polymerase used (P-value `r ifelse(max(filter(kruskal_chimera_prev, sub_sample_level > 500)[, "bh"]) < 0.05, invisible("<"), invisible("check data"))` 0.05) [Table S9]. Differences between Platinum and all other HiFi DNA polymerases accounted for the vast majority of these differences independent of cycle number and sub-sampling depth when using a Dunn's post-hoc test [Table S10]. Across sub-sampling depth and cycle number Accuprime^TM^ had the lowest chimera prevalence of all the HiFi DNA Polymerases regardless of whether 'pre.cluster' or chimera removal with VSEARCH had been used [Figure 5].

>>>>

***The Number of OTUs is Dependent on HiFi DNA Polymerase:***
There was a HiFi DNA polymerase dependent difference that was consistent across sub-sampling [Figure 1]. Lower cycle numbers (15-20) tended to result in less differences between HiFI DNA polymerase while cycle numbers of 25, 30, and 35 tended to have more distinct differences [Figure 1]. All sub-sampling levels were significantly different only at 35 cycles (P-value `r ifelse(max(filter(fs_anova_summary, cycle == "35x")[, "bh"]) < 0.0001, invisible("<"), invisible("check data"))` 0.0001) [Table S1]. There were differences in HiFi DNA polymerase at 25 and 30 cycles but the sub-sampling depth had to be 5000 or higher (P-value `r ifelse(max(filter(fs_anova_summary, cycle == "30x", sub_sample_level != 1000)[, "bh"]) < 0.05 & max(filter(fs_anova_summary, cycle == "25x", sub_sample_level != 1000)[, "bh"]) < 0.05, invisible("<"), invisible("check data"))` 0.05) [Table S1]. Platinum HiFi DNA polymerase was the main reason for the differences observed across all sub-sampling depths at 35 cycles, based on a Tukey post-hoc test (P-value `r ifelse(max(filter(fs_tukey_summary, padj < 0.05)[, "padj"]) < 0.05,  invisible("<"), invisible("check data"))` 0.05) [Table S2].

This HiFi DNA polymerase dependent difference in the number of OTUs was also observed in the mock community samples with the same DNA polymerases having high (Platinum) and low (Accuprime) number of OTUs [Figure 2 & Table S3]. Differences between HiFi DNA polymerase were observed as early as 20 cycles and a sub-sampling depth of 1000 sequences (P-value = `r format(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "bh"], digits = 1, nsmall = 1)`) [Table S3]. Using a Tukey post-hoc test differences between Platinum and the other HiFi DNA polymerases was the major driver of the differences seen at different cycle numbers and sub-sampling depths [Table S4]. Both fecal and mock samples consistently showed that across sub-sampling depth and cycle number the lowest number of OTUs identified was from Accuprime^TM^ while the highest was from Platinum [Figure 1 & 2]. Based on these observations it is clear that HiFi DNA polymerase affects the total number of OTUs observed within a sample.

***Minimal Bray-Curtis Differences are Detected by Cycle Number:***
For both fecal and mock samples, there was little difference between each respective 5-cycle increment at higher cycle numbers and this was consistent across the different sub-samplings [Figure 3]. There were only two differences between 5-cycle increments that were identified. First, there were differences for the same fecal sample between 20 versus 25 cycles that was independent of HiFi DNA polymerase but dependent on sub-sampling depth (sub-sampled to 1000 = `r round(median(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`) (median (25% - 75% quantile)), sub-sampled to 5000 = `r round(median(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`), sub-sampled to 10000 = `r round(median(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`)) [Figure 3A]. Second, where data is available for the mock communities, there were larger difference between 20 and 25 cycles (sub-sampled to 1000 = `r round(median(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`)) [Figure 3B]. However, these differences between the next 5-cycle increment do not persist once 25 cycles are reached [Figure 3].

Using PERMANOVA to test for differences within HiFi DNA polymerase groups based on cycle number, only Phusion had cycle dependent differences at 1000 and 5000 sub-sampling depth (P-value = `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 1000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)` and `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 5000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)`, respectively). Phusion was also one of only two HiFi DNA polymerases that that had samples for the 1000 sub-sampling depth at 15 cycles. Next, we assessed whether there were any major differences between 5-cycle increments within each sample. We found that there was no detectable difference in Bray-Curtis index when comparing to the previous 5-cycle increment (P-value `r ifelse(min(bray_sim_to_prev_cycle_kruskal_results$bh, na.rm = T) > 0.05, invisible(">"), invisible("check data"))` 0.05). However, Phusion at 1000 sub-sampling depth had a P-value = `r format(filter(bray_sim_to_prev_cycle_kruskal_results, sub_sample_level == 1000, taq == "PHU")[, "pvalue"], digits = 1, nsmall = 1)` before multiple comparison correction. It should be noted that at higher sub-sampling depths these differences in Bray-Curtis indices disappear for both differences in cycle number and within 5-cycle increments within an individual. Overall, HiFi DNA polymerase differences in Bray-Curtis index are dependent on both sub-sampling depth and cycle number.

***Sequence Error is Dependent on both HiFi DNA Polymerase and Cycle Number:***
Differences by HiFi DNA polymerase in the median average per base error varied without a clear pattern across sub-sampling depth [Table S5]. The highest per base median average error rates were for the Kappa HiFi DNA polymerase [Figure 4]. This error rate was minimally affected by both the 'pre.cluster' step and chimera removal by VSEARCH [Figure 4]. There were small differences in the per base error rate between the various HiFi DNA polymerase at lower cycle numbers and larger differences at higher cycle number with Platinum having the largest differences of all the HiFi DNA polymerase [Figure 4B-C and Table S6].

The total sequences with at least one error had multiple differences at different cycle numbers and was mostly alleviated by the use of the 'pre.cluster' step [Figure S1]. Major differences before this 'pre.cluster' step were driven by large differences in Accuprime^TM^ and Platinum versus the other HiFi DNA polymerase tested [Figure S1 & Table S7 & S8]. Although Accuprime^TM^ had the lowest per base error rate it had the largest number of sequences with at least one error, regardless of cycle number or sub-sampling depth [Figure S1]. However, this increased number of sequences with an error was lowered with existing bioinformatic approaches [Figure S1]. Investigation of whether there were HiFi DNA polymerase dependent effects on base substitution found that there were no biases in the types of substitution made [Figure S2]. Although sequence error is dependent on HiFi DNA polymerase it can be corrected using exsiting bioinformatic approaches.

***Chimeric Sequences are HiFi DNA Polymerase Dependent and Correlate with Number of OTUs:***
At all levels of sub-sampling and cycle number there were significant differences between the HiFi DNA polymerase used and chimera prevalence (P-value `r ifelse(max(filter(kruskal_chimera_prev, sub_sample_level > 500)[, "bh"]) < 0.05, invisible("<"), invisible("check data"))` 0.05) [Table S9]. Differences between Platinum and all other HiFi DNA polymerases accounted for the majority of these differences independent of cycle number and sub-sampling depth when using a Dunn's post-hoc test [Table S10]. Across sub-sampling depth and cycle number Accuprime^TM^ had the lowest chimera prevalence of all the HiFi DNA Polymerases regardless of whether 'pre.cluster' or chimera removal with VSEARCH was used [Figure 5].

### 2018-02-20 submission/manuscript.tex
update results section

\textbf{\emph{The Number of OTUs is Dependent on the HiFi DNA Polymerase
Used:}} In order to compare the number of OTUs across individuals a
Z-score normalization by individual, by cycle number, was employed on
the number of OTUs data. After normalization, we identified that there
was a HiFi DNA polymerase dependent difference that was consistent
across sub-sampling {[}Figure 1{]}. Lower cycle numbers (15-20) tended
to result in less differences between HiFI DNA polymerase while cycle
numbers of 25, 30, and 35 tended to have more distinct differences
{[}Figure 1{]}. All sub-sampling levels were significantly different
only at 35 cycles (P-value \textless{} 0.0001) {[}Table S1{]}. There
were differences in HiFi DNA polymerase at 25 and 30 cycles but the
sub-sampling depth had to be 5000 or higher (P-value \textless{} 0.05)
{[}Table S1{]}. Platinum HiFi DNA polymerase was the main driver of the
differences observed across all sub-sampling depths at 35 cycles, based
on a Tukey post-hoc test (P-value \textless{} 0.05) {[}Table S2{]}.

This HiFi DNA polymerase dependent difference in the number of OTUs was
also observed in the mock community samples with the same DNA
polymerases having high (Platinum) and low (Accuprime) number of OTUs
{[}Figure 2 \& Table S3{]}. Conversely, differences between HiFi DNA
polymerase were observed as early as 20 cycles and a sub-sampling depth
of 1000 sequences (P-value = 0.002) {[}Table S3{]}. Using a Tukey
post-hoc test differences between Platinum and the other HiFi DNA
polymerases was the major driver of the differences seen at different
cycle numbers and sub-sampling depths {[}Table S4{]}. Both fecal and
mock samples consistently showed that across sub-sampling depth and
cycle number the lowest number of OTUs identified was from
Accuprime\textsuperscript{TM} while the highest was from Platinum for
both fecal and mock samples {[}Figure 1 \& 2{]}.

\textbf{\emph{Minimal Bray-Curtis Differences are Detected by Cycle
Number:}} For both fecal and mock samples, there was little difference
between each respective 5-cycle increment at higher cycle numbers and
this was consistent across the different sub-samplings {[}Figure 3{]}.
Overall, there were only two differences between 5-cycle increments that
were identified. First, there were differences for the same fecal sample
between 20 versus 25 cycles that was independent of HiFi DNA polymerase
but dependent on sub-sampling depth (sub-sampled to 1000 = 0.51 (0.4 -
0.79) (median (25\% - 75\% quantile)), sub-sampled to 5000 = 0.43 (0.33
- 0.63), sub-sampled to 10000 = 0.4 (0.24 - 0.43)) {[}Figure 3A{]}.
Second, for the mocks, where data is available, there were larger
difference between 20 and 25 cycles (sub-sampled to 1000 = 0.88 (0.42 -
0.91)) {[}Figure 3B{]}. However, these differences between the next
5-cycle increment do not persist once 25 cycles are reached {[}Figure
3{]}.

Using PERMANOVA to test for differences within HiFi DNA polymerase
groups based on cycle number, only Phusion had cycle dependent
differences at 1000 and 5000 sub-sampling depth (P-value = 0.03 and
0.01, respectively). Phusion was also one of only two HiFi DNA
polymerases that managed to have samples for the 1000 sub-sampling depth
at 15 cycles. Next, we assessed whether there were any major differences
between 5-cycle increments within each sample. We found that there was
no detectable difference in Bray-Curtis index when comparing to the
previous 5-cycle increment (P-value \textgreater{} 0.05). However,
Phusion at 1000 sub-sampling depth had a P-value = 0.02 before multiple
comparison correction. It should be noted that at higher sub-sampling
depths these differences in Bray-Curtis indices disappear for both
differences in cycle number and within 5-cycle increments within an
individual.

\textbf{\emph{Sequence Error is Dependent on both HiFi DNA Polymerase
and Cycle Number:}} Differences by HiFi DNA polymerase in the median
average per base error varied without a clear pattern across
sub-sampling depth {[}Table S5{]}. Generally, the highest per base
median average error rates were for the Kappa HiFi DNA polymerase
{[}Figure 4{]}. This error rate was minimally affected by both the
`pre.cluster' step and chimera removal by VSEARCH {[}Figure 4{]}. There
were small differences in the per base error rate between the various
HiFi DNA polymerase at lower cycle numbers and larger differences at
higher cycle number with Platinum having the largest differences of all
the HiFi DNA polymerase {[}Figure 4B-C and Table S6{]}.

The total sequences with at least one error had multiple differences at
different cycle numbers and was mostly alleviated by the use of the
`pre.cluster' step {[}Figure S1{]}. Major differences before this
`pre.cluster' step were driven by large differences in
Accuprime\textsuperscript{TM} and Platinum versus the other HiFi DNA
polymerase tested {[}Figure S1 \& Table S7 \& S8{]}. Although
Accuprime\textsuperscript{TM} had the lowest per base error rate it had
the largest number of sequences with at least one error, regardless of
cycle number or sub-sampling depth {[}Figure S1{]}. However, this
increased number of sequences with an error can be drastically lowered
with existing bioinformatic approaches {[}Figure S1{]}. Investigation of
whether there were HiFi DNA polymerase dependent effects on base
substitution found that there were generally no biases in the types of
substitution made {[}Figure S2{]}.

\textbf{\emph{Chimeric Sequences are HiFi DNA Polymerase Dependent and
Correlate with Number of OTUs:}} After chimera removal using VSEARCH and
removal of sequences that did not classify as bacteria, we assessed the
percentage of sequences that were still chimeric within our mock
community. At all levels of sub-sampling and cycle number there were
significant differences between the HiFi DNA polymerase used (P-value
\textless{} 0.05) {[}Table S9{]}. Differences between Platinum and all
other HiFi DNA polymerases accounted for the vast majority of these
differences independent of cycle number and sub-sampling depth when
using a Dunn's post-hoc test {[}Table S10{]}. Across sub-sampling depth
and cycle number Accuprime\textsuperscript{TM} had the lowest chimera
prevalence of all the HiFi DNA Polymerases regardless of whether
`pre.cluster' or chimera removal with VSEARCH had been used {[}Figure
5{]}.

>>>>

\textbf{\emph{The Number of OTUs is Dependent on HiFi DNA Polymerase:}}
There was a HiFi DNA polymerase dependent difference that was consistent
across sub-sampling {[}Figure 1{]}. Lower cycle numbers (15-20) tended
to result in less differences between HiFI DNA polymerase while cycle
numbers of 25, 30, and 35 tended to have more distinct differences
{[}Figure 1{]}. All sub-sampling levels were significantly different
only at 35 cycles (P-value \textless{} 0.0001) {[}Table S1{]}. There
were differences in HiFi DNA polymerase at 25 and 30 cycles but the
sub-sampling depth had to be 5000 or higher (P-value \textless{} 0.05)
{[}Table S1{]}. Platinum HiFi DNA polymerase was the main reason for the
differences observed across all sub-sampling depths at 35 cycles, based
on a Tukey post-hoc test (P-value \textless{} 0.05) {[}Table S2{]}.

This HiFi DNA polymerase dependent difference in the number of OTUs was
also observed in the mock community samples with the same DNA
polymerases having high (Platinum) and low (Accuprime) number of OTUs
{[}Figure 2 \& Table S3{]}. Differences between HiFi DNA polymerase were
observed as early as 20 cycles and a sub-sampling depth of 1000
sequences (P-value = 0.002) {[}Table S3{]}. Using a Tukey post-hoc test
differences between Platinum and the other HiFi DNA polymerases was the
major driver of the differences seen at different cycle numbers and
sub-sampling depths {[}Table S4{]}. Both fecal and mock samples
consistently showed that across sub-sampling depth and cycle number the
lowest number of OTUs identified was from Accuprime\textsuperscript{TM}
while the highest was from Platinum {[}Figure 1 \& 2{]}. Based on these
observations it is clear that HiFi DNA polymerase affects the total
number of OTUs observed within a sample.

\textbf{\emph{Minimal Bray-Curtis Differences are Detected by Cycle
Number:}} For both fecal and mock samples, there was little difference
between each respective 5-cycle increment at higher cycle numbers and
this was consistent across the different sub-samplings {[}Figure 3{]}.
There were only two differences between 5-cycle increments that were
identified. First, there were differences for the same fecal sample
between 20 versus 25 cycles that was independent of HiFi DNA polymerase
but dependent on sub-sampling depth (sub-sampled to 1000 = 0.51 (0.4 -
0.79) (median (25\% - 75\% quantile)), sub-sampled to 5000 = 0.43 (0.33
- 0.63), sub-sampled to 10000 = 0.4 (0.24 - 0.43)) {[}Figure 3A{]}.
Second, where data is available for the mock communities, there were
larger difference between 20 and 25 cycles (sub-sampled to 1000 = 0.88
(0.42 - 0.91)) {[}Figure 3B{]}. However, these differences between the
next 5-cycle increment do not persist once 25 cycles are reached
{[}Figure 3{]}.

Using PERMANOVA to test for differences within HiFi DNA polymerase
groups based on cycle number, only Phusion had cycle dependent
differences at 1000 and 5000 sub-sampling depth (P-value = 0.03 and
0.01, respectively). Phusion was also one of only two HiFi DNA
polymerases that that had samples for the 1000 sub-sampling depth at 15
cycles. Next, we assessed whether there were any major differences
between 5-cycle increments within each sample. We found that there was
no detectable difference in Bray-Curtis index when comparing to the
previous 5-cycle increment (P-value \textgreater{} 0.05). However,
Phusion at 1000 sub-sampling depth had a P-value = 0.02 before multiple
comparison correction. It should be noted that at higher sub-sampling
depths these differences in Bray-Curtis indices disappear for both
differences in cycle number and within 5-cycle increments within an
individual. Overall, HiFi DNA polymerase differences in Bray-Curtis
index are dependent on both sub-sampling depth and cycle number.

\textbf{\emph{Sequence Error is Dependent on both HiFi DNA Polymerase
and Cycle Number:}} Differences by HiFi DNA polymerase in the median
average per base error varied without a clear pattern across
sub-sampling depth {[}Table S5{]}. The highest per base median average
error rates were for the Kappa HiFi DNA polymerase {[}Figure 4{]}. This
error rate was minimally affected by both the `pre.cluster' step and
chimera removal by VSEARCH {[}Figure 4{]}. There were small differences
in the per base error rate between the various HiFi DNA polymerase at
lower cycle numbers and larger differences at higher cycle number with
Platinum having the largest differences of all the HiFi DNA polymerase
{[}Figure 4B-C and Table S6{]}.

The total sequences with at least one error had multiple differences at
different cycle numbers and was mostly alleviated by the use of the
`pre.cluster' step {[}Figure S1{]}. Major differences before this
`pre.cluster' step were driven by large differences in
Accuprime\textsuperscript{TM} and Platinum versus the other HiFi DNA
polymerase tested {[}Figure S1 \& Table S7 \& S8{]}. Although
Accuprime\textsuperscript{TM} had the lowest per base error rate it had
the largest number of sequences with at least one error, regardless of
cycle number or sub-sampling depth {[}Figure S1{]}. However, this
increased number of sequences with an error was lowered with existing
bioinformatic approaches {[}Figure S1{]}. Investigation of whether there
were HiFi DNA polymerase dependent effects on base substitution found
that there were no biases in the types of substitution made {[}Figure
S2{]}. Although sequence error is dependent on HiFi DNA polymerase it
can be corrected using exsiting bioinformatic approaches.

\textbf{\emph{Chimeric Sequences are HiFi DNA Polymerase Dependent and
Correlate with Number of OTUs:}} At all levels of sub-sampling and cycle
number there were significant differences between the HiFi DNA
polymerase used and chimera prevalence (P-value \textless{} 0.05)
{[}Table S9{]}. Differences between Platinum and all other HiFi DNA
polymerases accounted for the majority of these differences independent
of cycle number and sub-sampling depth when using a Dunn's post-hoc test
{[}Table S10{]}. Across sub-sampling depth and cycle number
Accuprime\textsuperscript{TM} had the lowest chimera prevalence of all
the HiFi DNA Polymerases regardless of whether `pre.cluster' or chimera
removal with VSEARCH was used {[}Figure 5{]}.
